Topological approach to finding blackholes in directed networks

### 2019-12-08 blackholes/pavt/article.tex
Small text changes + algo updated

\FIXME{не совсем понял логику в общем тексте}
Of course, we could straightforward build the reachability matrix. But, it takes $O(V^3)$ time. Even though we decreased graph size
in preprocessing step, it still can result in lots of computation. It is worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
Difficult precalculation will affect the time of first blackhole discovery. That is why, we will only try to partially skip duplicates and the rest will
be filtered later in lazy manner.

@@

@@

@@

Let's consider a closure of a single global root $r$. Let this closure to be acquired by depth-first search, which basically means, that
we can iterate by $i$ over these nodes in topologically sorted order $topsortOrder_r[i]$.

Also, while traversing this graph, we can calculate size of closure for each node
in the given root closure. Finally, if some node is in the $i$-th position of the topological sort and its closure size is $i$, then we can 
conclude, that every node with index less than $i$ in the topological sort will be reachable from $topsortOrder_r[i]$. We will call such nodes special.

If there are two special nodes under the same root, union of their closures will be equal to the closure of topologically higher node.
Hence, union of closures for any subset of special nodes under the same root will result into closure of the highest special node.

%TODO: Осознать до конца
Let's consider two nodes in the closure of a global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
the special one should be topologically lower than non-special. Otherwise it would be one node basis.
To summarize, if given a set of nodes, we shall remove nodes topologically lower than highest ones special nodes in every global root closure.

@@

Algorithm \ref{alg:isbasis} IsBasis tests a set of nodes as a blackhole basis. 
First, we need to know what special nodes there are. This step can be precalculated. 
\FIXME{как вычисляются global roots?}
To find out, we take every global root of the condensated graph and find topological sort of its closure.
Those vertices, which satisfy the rule above, we mark as special.

\FIXME{зачем второй этап?}
Second, we calculate how many special nodes for every root there is in the given set. If there is more than one special node for root, the $Cand$ set is called non-basis and will be omitted in the high-level TopSort blackhole mining Algorithm \ref{alg:topsort}.

%\FIXME{partially implemented: по факту тут алгоритм применяется только к первому корню. Это было проще реализовать. }

Also we check if there are any non-special nodes lower than the special ones in the related root topological order, and if those found, then omit the node set.

Third, when we are out of heuristics, we check if any of the candidate nodes are reachable from one another. If so, we omit this set.

@@

\begin{figure}[H]
	\begin{center}
		\begin{algorithm}[H]
			\SetAlgoLined
			\SetKwInOut{Input}{Input}
			\SetKwInOut{Output}{Output}
			\Input{$G(V,E)$ -- directed acyclic graph (after graph condensation) \\
                                $Cand \subset V$ -- set of nodes, candidate to be basis}
			\Output{$True, if\ Cand\ is\ a\ basis, False\ otherwise$}
                        \tcc{(1) calculate special nodes for every root} 
                        \ForEach{$r \in Roots(G)$} {
                            $Special_r = \emptyset$ \\
                            $build\ topsortOrder_r$ \\
                            \For{$i = 0; i<|topsortOrder_r|; i=i + 1$} {
                                $v = topsortOrder_r[i]$ \\
                                $C_v = Closure(v)$ \\
                                \If{$|C_v| == i$} {
                                    $Special_r = Special_r \cup v$ \\
                                }
                            }
                        }
			\tcc{(2) check for excluding specials} 
			\ForEach{$v \in Cand$} {
                            \ForEach{$r \in Roots(G)$} {
                                \If{$v \in Special_r$} {
                                    $Spec_r = Spec_r \cup v$ \\
                                    \If{$|Spec_r| > 1$} {
                                        \Return $False$ \\
                                    }
                                }
                            }
			}
                        \tcc{(3) check for candidates reachable from one another}
                        \ForEach{$(v, u) \in Cand(G) \times Cand(G)$} {
                            \If{v is reachable from u} {
                                \Return $False$ \\
                            }
                        }
                        \Return $True$
			\label{alg:isbasis}
			\caption{IsBasis}
		\end{algorithm}
	\end{center}
\end{figure}

>>>>

\FIXME{не совсем понял логику в общем тексте}
Of course, we could straightforward build the reachability matrix. But, it takes \FIXME{ Not sure $O(V^3)$ } time. Even though we decreased graph size
in preprocessing step, it still can result in lots of computation. It is worth mentioning that ultimate goal is to find at least some blackholes as soon as possible. 
Difficult precalculation will affect the time of first blackhole discovery. That is why, we will only try to partially skip duplicates and the rest will
be filtered later in lazy manner.

@@

\begin{remark}
    Specialness of a node is decided independently under different roots. One node can be special for root $r1$ and non special for $r2$ at
    the same time.
\end{remark}

@@

\begin{remark}
    If given two special nodes under the same root, ont of them is always dominated by the other.
\end{remark}

@@

%TODO: Осознать до конца
Let's consider two nodes in the closure of a global root. One of them is non-special, the other one is special. In order for these two nodes to be basis
the special one should not dominate non-special. Otherwise non-special node can be omitted from basis.
In general, if given a set of nodes, we shall remove all dominated ones in order for the rest to be a proper basis.

@@

Algorithm \ref{alg:isbasis} IsBasis tests a set of nodes as a blackhole basis. 
First, we need to know what special nodes there are. This step can be precalculated. 
\FIXME{как вычисляются global roots?}
To find out, we take every global root of the condensated graph and find topological sort of its closure.
Those vertices, which satisfy definition, we mark as special for according roots.

\FIXME{зачем второй этап?}
%\FIXME{partially implemented: по факту тут алгоритм применяется только к первому корню. Это было проще реализовать. }

Second, if there are any dominated nodes (special or non-special) under any of the roots, we omit the node set.

Third, when we are out of cheap heuristics, we check if any of the candidate nodes are reachable from one another. If so, we omit this set.

@@

\begin{figure}[H]
	\begin{center}
		\begin{algorithm}[H]
			\SetAlgoLined
			\SetKwInOut{Input}{Input}
			\SetKwInOut{Output}{Output}
			\Input{$G(V,E)$ -- directed acyclic graph (after graph condensation) \\
                                $Cand \subset V$ -- set of nodes, candidate to be basis}
			\Output{$True, if\ Cand\ is\ a\ basis, False\ otherwise$}
                        \tcc{(1) calculate special nodes for every root} 
                        \ForEach{$r \in Roots(G)$} {
                            $Special_r = \emptyset$ \\
                            $build\ topSortOrder_r$ \\
                            \For{$i = 0; i<|topSortOrder_r|; i=i + 1$} {
                                $v = topSortOrder_r[i]$ \\
                                $C_v = Closure(v)$ \\
                                \If{$|C_v| == i$} {
                                    $Special_r = Special_r \cup v$ \\
                                    $maxSpecialIndex_r = max(maxSpecialIndex_r, i)$ \\
                                }
                            }
                        }
			\tcc{(2) check for dominated nodes} 
                        \ForEach{$r \in Roots(G)$} {
                            $maxSpecialIndex_r = -1$ \\
                        }
			\ForEach{$v \in Cand$} {
                            \ForEach{$r \in Roots(G)$} {
                                \If{$v \in Special_r$} {
                                    $vTopSortIndex_r = arg_v (topSortOrder_r)$ \\
                                    $maxSpecialIndex_r = max(maxSpecialIndex_r, vTopSortIndex_r)$ \\
                                }
                            }
			}
			\ForEach{$v \in Cand$} {
                            \ForEach{$r \in Roots(G)$} {
                                \If{$v \in Closure(r)$} {
                                    $vTopSortIndex_r = arg_v (topSortOrder_r)$ \\
                                    \If{$vTopSortIndex_r < maxSpecialIndex_r$} {
                                        \Return $False$ \\
                                    }
                                }
                            }
			}
                        \tcc{(3) check for candidates reachable from one another}
                        \ForEach{$(v, u) \in Cand(G) \times Cand(G)$} {
                            \If{v is reachable from u} {
                                \Return $False$ \\
                            }
                        }
                        \Return $True$
			\label{alg:isbasis}
			\caption{IsBasis}
		\end{algorithm}
	\end{center}
\end{figure}
